\documentclass[12pt]{article}

\usepackage{tikz} % картинки в tikz
\usepackage{microtype} % свешивание пунктуации

\usepackage{array} % для столбцов фиксированной ширины

\usepackage{indentfirst} % отступ в первом параграфе

\usepackage{sectsty} % для центрирования названий частей
\allsectionsfont{\centering}

\usepackage{amsmath, amssymb, amsthm} % куча стандартных математических плюшек

\usepackage{amsfonts}

\usepackage{comment}

\usepackage[top=2cm, left=1.2cm, right=1.2cm, bottom=2cm]{geometry} % размер текста на странице

\usepackage{lastpage} % чтобы узнать номер последней страницы

\usepackage{enumitem} % дополнительные плюшки для списков
%  например \begin{enumerate}[resume] позволяет продолжить нумерацию в новом списке
\usepackage{caption}


\usepackage{hyperref} % гиперссылки

\usepackage{multicol} % текст в несколько столбцов


\usepackage{fancyhdr} % весёлые колонтитулы
\pagestyle{fancy}
\lhead{Stochastic Calculus}
\chead{Final exam 2020}
\rhead{2020-12-29}
\lfoot{}
\cfoot{DON'T PANIC}
\rfoot{}
\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0.4pt}

\let\P\relax
\DeclareMathOperator{\P}{\mathbb{P}}
\DeclareMathOperator{\Cov}{\mathbb{C}ov}
\DeclareMathOperator{\E}{\mathbb{E}}
\DeclareMathOperator{\Var}{\mathbb{V}ar}
\newcommand{\cN}{\mathcal{N}}

\usepackage{todonotes} % для вставки в документ заметок о том, что осталось сделать
% \todo{Здесь надо коэффициенты исправить}
% \missingfigure{Здесь будет Последний день Помпеи}
% \listoftodos - печатает все поставленные \todo'шки


% более красивые таблицы
\usepackage{booktabs}
% заповеди из докупентации:
% 1. Не используйте вертикальные линни
% 2. Не используйте двойные линии
% 3. Единицы измерения - в шапку таблицы
% 4. Не сокращайте .1 вместо 0.1
% 5. Повторяющееся значение повторяйте, а не говорите "то же"



\usepackage{fontspec}
\usepackage{polyglossia}

\setmainlanguage{english}
\setotherlanguages{russian}

% download "Linux Libertine" fonts:
% http://www.linuxlibertine.org/index.php?id=91&L=1
\setmainfont{Linux Libertine O} % or Helvetica, Arial, Cambria
% why do we need \newfontfamily:
% http://tex.stackexchange.com/questions/91507/
\newfontfamily{\cyrillicfonttt}{Linux Libertine O}

\AddEnumerateCounter{\asbuk}{\russian@alph}{щ} % для списков с русскими буквами
% \setlist[enumerate, 2]{label=\asbuk*),ref=\asbuk*}




\begin{document}


Short rules: 120 minutes, A4 cheat sheet and calculator are allowed.


Copy and sign this short pledge:

\vspace{10pt}
\textit{«I pledge on my honor that I will not give or receive any unauthorized 
assistance on this exam».}
\vspace{10pt}


And now you may solve the problems!


\begin{enumerate}

    \item (10 points) Ded Moroz would like to receive $K$ roubles at time $T=2$ if $S_2< 0.5S_1$ and zero otherwise.
    
    Assume the framework of Black and Scholes model, $S_t$ is the share price, $r$ is the risk free rate,
    $\sigma$ is the volatility. 

    How much Ded Moroz should pay now at $t=0$?
    
    \item (10 points) Simplify as much as possible the integral
    \[
    \int_0^t \exp(-W_u - u/2) dW_u.    
    \]

    \item (10 points) The process $Y_t$ is defined by 
    \[
    dY_t = W_t^2 dt  + W_t dW_t, \; Y_0 = 0.    
    \]
    \begin{enumerate}
        \item (6 points) Find $\E(Y_t)$, $\E(Y_t W_t)$, $\E(Y_t W_t^2)$.
        \item (4 points) Find $\Var(Y_t)$.
    \end{enumerate}

    \item (10 points) Find at least one solution of the stochastic differential equation
    \[
    dR_t = 4R_t dt + 7 dW_t, \quad R_0 = 1.
    \]

    You are free to use the following steps:
    \begin{enumerate}
        \item Solve deterministic equation $dQ_t = 4Q_t dt$.
        \item Now you need to «remove» this deterministic solution $Q_t$ from $R_t$. 
        To accomplish this goal represent $R_t$ as $R_t = Q_t B_t$ and find a very simple equation for $B_t$.
        \item Solve the equation for $B_t$.
        \item Finalize your solution.
    \end{enumerate}

    \item (10 points) The variables $X$ and $Y$ are independent and are exponentially distributed with mean 1.
    Let $L = \min \{X, Y\}$ and $R = \max \{X, Y\}$. 

    Find $\E(L \mid R)$ and $\E(R\mid L)$.

    \vspace{20pt}

    Be brave!! One more problems is waiting to kill you!! The final boss of this game!!

    \newpage


    \item (20 points) Consider a process $S_t = Z_1 + Z_2 + \ldots + Z_t$ with $S_0=0$.
    Increments $Z_t$ are identically distributed and take values $(-1)$ and $1$ 
    with equal probabilities. 

    Let $\tau$ be the first moment of time when $|S_t| = 100$.
    Let's denote by $Z$ the indicator that $S_\tau = 100$.

    \begin{enumerate}
        \item Find $f(\lambda)$ such that $M_t = f(\lambda)^t\exp(\lambda S_t)$ is a martingale.
        \item Using Doob's optional stopping time theorem obtain
        an equation for $\E(Zf(\lambda)^\tau)$ and $\E((1-Z)f(\lambda)^\tau)$.
        \item Solve for these two expected values.
        \item Find $\E(f(\lambda)^\tau)$.
        \item Find $h(\alpha) = \E(\exp(\alpha \tau))$ for arbitrary $\alpha$.
        \item Find $h'(0)$, $h''(0)$ and hence $\E(\tau^2)$.
    \end{enumerate}

    Hint: you may not check technical conditions of Doob's optional stopping time theorem.
    
\end{enumerate}


\end{document}
